---
title: "R basics"
author: "Sophia Barrett"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R

## Setup

By now, you should have R and Rstudio installed on your local machine (that is, your own computer).  If not follow the steps below:

1. Visit the CRAN R website at https://cran.r-project.org/ and install the version for R that is appropriate for your machine.  

2.  AFTER YOU HAVE INSTALLED R, you need to visit the RStudio downloads page at https://rstudio.com/products/rstudio/download/ and download the FREE Rstudio Desktop.

While you are perfectly able to use R, Rstudio provides a nice user-friendly "face" for the program.  Make sure to watch the video for a introduction to RStudio.

## R basics
### R as a calculator

First, you can use R as a calculator. So we use `+` for addition:
```{r}
2 + 3
```
We use `-` for subtraction:
```{r}
12-30
```

We use `*` for multiplication: 
```{r}
60*2
```

We use `/` for division:
```{r}
10/3
```
We use `**` for exponentiation: 
```{r}
10**2
```
We can also use `^` for exponents:
```{r}
12 ^ 2
```

We have additional functions for `integer division` and `remainder division`.

Let's say we want to calculate 25 divided by 4:
```{r}
25/4
```

Integer division is just getting the whole number portion of the quotient (the number before the decimal point); in this case, that would be 6.

We use `%/%` for integer division:
```{r}
25 %/% 4
```

Now let's think about the remainder.  4 goes into 25 six times (that is our integer part).  Since 4 times 6 equals 24, we have 1 left over (because 25 - 24 = 1 --> this is our remainder).  We have a remainder of 1 as a result of dividing 25 by 4.

Remainder division is better known as `modulo division` and the remainder is the `modulus`.

We use `%%` for remainder division:
```{r}
25 %% 4
```

Note that R will follow the order of operatons.  So, 
```{r}
8 / 2 * (2 + 2)
```
This problem always trips up people.  Make sure you understand why 16 is the correct answer.

### Assignment to variable

In the examples above, we had R make a calculation and then result showed directly to screen.  

We can also store the result in a variable for later use.  For example:

```{r}
x <-  2 + 3
```
`<-` is R's assignment operator.  It works like `=` and, while = will work, the practice is to use the assignment operator.

When you run the code, nothing comes to screen.  Why? Becuase we told R to store the value of 2 + 3 in the variable x.  If we want to see what is stored in x, we have to call x by typing its name:

```{r}
x
```

We can also use x in equations (like you did in algebra class).

```{r}
x + 2
```

```{r}
x * x
```

Now is a good time to point out that R is **case sensitive** so x is not the same as X.  If you get an error, the name of the variable is a good starting place.  Check the spelling and check the case.

### Variable types in R

As we can see, we can use numbers to work in R as numbers are part of the `numeric` variable type.  There are two other variable types we will use in this class:  `character` and `logical`.  

As we saw, `numerical variables` are numbers on which we can do mathematical operations.

```{r}
a <- 2^10 + 6
```
```{r}
a
```

`Charater variables` are `strings`. Strings are a sequence of letters, numbers, and symbols.  To create a string, use **double quotes**.

```{r}
message <- "Hello World"
```

```{r}
message
```

`Logical variables` are also called `Boolean variables`.  Logical variables can on two values: TRUE or FALSE.  **Remember that R is case-sensitive so you have to write the words TRUE and FALSE in all capital letters**.  However, you can also write T or F, respectively.

```{r}
y <-  FALSE
```


```{r}
y
```


```{r}
renter <- T
```


```{r}
renter
```

For boolean values, FALSE is equivalent to O and TRUE is equivalent to 1.  This make it possible to do Boolean algebra.  For example,

```{r}
m <- TRUE
n <- FALSE
```

Note: we can have more than one command in notebook cell or code chunk.

```{r}
print(m * 1)
print(n * 1)
```
The `print()` is a built-in R function that allows us to print messages or calcuations to the screen.

For example:
```{r}
print("It is a lovely day!")
```

What you put between the parentheses is called the `argument` or the `parameter`.  In this example, the message to be printed is the argument.  You place string messages in double quotes.  

You can also print numerical or logical values.

```{r}
print(2 + 3)
```
Most times, we will not have to use the print statement, as you have seen in our examples.

One more thing to note, R will **overwrite** your existing variables without a warning.

```{r}
firstname = "Julie"
```


```{r}
firstname
```


```{r}
firstname = "Sophie"
```


```{r}
firstname
```
```{r}
# Name
```
Note the error!  I used 'Name' (capital N), instead of 'name' (lowercase n and the way I actually created the variable).  Remember R is case sensitive so name and Name are different variables.

One final note, we can use a hashtag `#` to `comment out a line`.  This means that anything after the hashtag is ignored by R.  It is a great way to make notes to ourselves about the code.

```{r}
# This line is a comment and will be 
#ignored by R
```

**Remember, R is _case-sensitive_ **

