---
title: "2_Vectors"
author: "Sophia Barrett"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vectors

## Data Structure
A **data structure** is a way of organizing and storing data.  There are different types of data structures in R.  The first one we will look will be **vectors**. 

## Our first data structure: Vectors

Vectors are one of the basic data structures in R.  A vector is a sequence of elements of the same data type (numeric, character, or logical).  Vectors are one-dimensional, meaning they just have a length.

## Getting help

You can get help in R in 3 ways  
1.  Use the `help()` command  
2.  Use the question mark  
3.  Google it!  

From the help command window you can read more about functions.  

```{r}
help(mean)
```

```{r}
?c
```

## Making functions with `c()`
`c()` stands for concatenate (or combine).  Using the c() function is the most common way to make vectors of any data type.

Make a vector of values from 1 to 9:
```{r}
c(1, 2, 3, 4, 5, 6, 7, 8, 9)
```
In the example above, we made a vector but did not save it to a variable for later use.  Let's save it to the variable x.

```{r}
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
```

```{r}
x
```
The elements of a vector are called compents; the vector x is made up of numerical components from 1 to 9.

```{r}
q <-  c(10, 8, 6, 4, 2, 0, -2, -4, -6, -8, -10)
q
```

In addition to numeric vectors, we can create character vectors and logical vectors.

```{r}
months <- c("April", "May", "June", "July", "August")
months
```

```{r}
result <- c(TRUE, FALSE, T, T, T, F, F, FALSE, TRUE)
result
```
### Getting the length of a vector
To get the length of a vector means to get the number of components in the vector.  We use the `length()` fucntion to determine the length of a vector.

```{r}
length(x)
```
```{r}
length(months)
```


```{r}
length(result)
```

### Coercion Rules
Remember, all the elements of a vector need to be of the same data type.  If you attempt to make a vector of different types, R will attempt to **coerce** the variable into a type that matches the rest of the elements.  

We can figure out the date type of a vector by using the `typeof()` function.

```{r}
test <- c(1, 2, 3)
typeof(test)
```
We see the test variable is a special kind of numeric vector called `double`. Double refers to decimal numbers, real numbers, and floating point values.

In the next example, the variable `mix` is a vector of three elements -- 2 numerical and 1 string.  What type of vector is mix?

```{r}
mix <- c(1, 2, "3")
typeof(mix)
```
we see that mix is a character vector.  In other words, R coerced (converted) the 1 and the 2 into string values. So it made mix into a character vector.
```{r}
mix
```
So when there is a mixture of numerical elements and character elements, R will coerce all the elements into strings.

What happens if you have a mixture of numerical elements and logical elements?

```{r}
hmmm <- c(1, 2, FALSE)
typeof(hmmm)
```

```{r}
hmmm
```

### Adding elements to an existing vector

```{r}
d <- c(10, 20, 30)
d
```

```{r}
d <- c(d, 50)
d
```

## Using the colon (:) operator to make vectors

We can also create vectors using a colon.  The resulting vector will be a **sequence** of **integers**, another type of number.  

```{r}
m <-  1:10    #note: no c() function
m
```
```{r}
typeof(m)
```


```{r}
n <- 100: 75
n
```


```{r}
p <- c(40:50, 100)
p
```
```{r}
typeof(p)
```

```{r}
i <- 5
j <- 15

z <- i:j
z
```


```{r}
y <- j:i
y
```


```{r}
```





































