---
title: "3_More_Vectors"
author: "Sophia Barrett"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# More Vectors

## The `rep()` function

We have other ways to make vectors.  For example, the `rep()` function allows us to make a vector of repeated values.  Here we will repeat the value of 10 three times:
```{r}
rep(10, 3)
```
We can also repeat string vectors: 
```{r}
rep("abc", 6)
```

and booleans: 
```{r}
rep(T, 10)
```

### A note about `str()`

We can use the structure function, `str()` to look at the strucure of an object
```{r}
m <- rep(45, 10)
str(m)
```
Here we see that m is a numeric vector of 10 elements.  We also see the value of the elements in vector m.

## Vector-wise Replication vs Element-wise Replication

We can also chose to repeat entire vectors or to repeat the elements of a vector.  If we choose to repeat the entire vector, this is called **vector-wise replication**.  In the following example, we will repeat the vector 10 times
```{r}
n <- rep(c(20, 30, 40), 10)
n
```

or we can choose to repeat each element of the vector with an **element-wise replication**, for example:
```{r}
p <- rep(c(20, 30, 40), each = 10)
p
```
By using the `each = ` parameter, we are telling R to repeat each element 10 times in this example. You will see this will come in handy when we look at differences between groups.  For example,

```{r}
gender = rep(c("Female", "Male"), 5)
gender
```

```{r}
gender2 = rep(c("Female", "Male"), each = 5)
gender2
```
What differences do you note?

### Using `rep()` with the colon

We can also use `rep()` with the colon.  For vector-wise replication
```{r}
q <- rep(50:60, 5)
q
```

For element-wise replication
```{r}
r <- rep(50:60, each = 5)
r
```

## Creating sequences of numbers with `seq()`

We can also create sequences of numbers using the sequence function, `seq()`.  To use the seq() function, we need to specify  
1.  a **start** value  
2.  a **stop** value  
3.  a **step**  

We can get help on the seq() by typing `help(seq)` or `?seq` to bring up documentation on the function.

```{r}
?seq
```

From the documentation, we can see the default values for the seq():   
the default starting value is 1(`from = 1`)  
the default ending value is 1 (`to = 1`)  
the default step value is 1 (`by = 1`)  


```{r}
seq()
```
This is not very useful so let's change the parameters

```{r}
seq(from = 0, to = 12, by = 2)
```


```{r}
seq(from = 1, to = 2, by = .10)
```

We need to be aware of the order of the arguments in R.  We can omit the labels **but** if we do not enter the arguments in the correct order, we will end up with the wrong result.  As such, you have to enter the argument in this exact order -- start, stop, step.

To create a sequence of values that start at 100, end at 1000, and go up in steps of 100, we can do:
```{r}
seq(100, 1000, 100)
```

To create a sequence of values that start at 70, ends at 50, and **decreases** in steps of 5:
```{r}
seq(70, 50, -5)
```

What do we get with the following:
```{r}
d <- seq(100, -100, -10)
d
```

Instead of telling R the steps it should increase or decrease by, we can tell it how many components we want instead.  In this case, R will then calculate the correct step.

To indicate the number of components, we use the `length = ` argument (we must use this label)

Let's create a sequence of n = 10 numbers between 20 and 60:
```{r}
f <- seq(20, 60, length = 10)
f
```

We can combine the rep() and the seq() functions in creating vectors:
```{r}
d <- rep(seq(1, 5, length = 3), each = 3)
d
```

## Creating Vectors of Random Numbers

### Using `sample()` to generate discrete random numbers

We can create a vector of discrete random numbers using the `sample()` function.  To use the sample() function, we need to indicate the population from which we are sampling and the number of values to be generated (i.e. the sample size, n).

Let's generate n = 15 discrete random numbers (integers) from 1 to 100:
```{r}
i <- sample(1:100, 15)
i
```

If we are starting from 1, we can also use the following code:
```{r}
j <- sample(100, 15)
j
```
Note, we get a different sample of n = 15 numbers since we are getting a **random** sample.


The population can also be a vector:
```{r}
k <- 250: 500   #k is a vector of integers from 250 to 500
l <- sample(k, 10)  #we will use the sample function to pick 10 random discrete numbers from 250 to 500
l
```

Review the random samples made for i, j, and l -- do any of the values repeat?

We should have no repeating values (i.e. we should have each vector being a vector of **unique** values).  This is becuase the samples are made **without replacement** -- if we do not replace the value, we cannot use it again.  

If you look at the documentation for sample(), we see that the `replace = ` parameter is set to FALSE by default.  This means R does sampling without replacement by default

```{r}
?sample
```
If we want to sample with replacement, we have to set the `replace = ` paramter to TRUE.

(We will revisit the idea of sampling with replacement when we get to probability and scores.)

If we want to generate n = 20 random numbers from 1 to 10 using sampling with replacement:
```{r}
m <- sample(10, 20, replace = TRUE)
m
```
**NOTE**: by if your n is better than the population you are pulling from, then you MUST use `replace = TRUE` as you will not have enough unique values to make your sample.


### Using `rnorm()` to generate normally-distributed random numbers

We can use the `rnorm()` function to generate a vector normally distributed random numbers.  By default, $\mu$ = 0 and standard deviation $\sigma$ = 1.  Using a mean of $\mu$ = 0 and standard deviation $\sigma$ = 1 gives us the standard normal distribution.

The **template** looks like:  rnorm(n, $\mu$, $\sigma$)

To get 100 normally distributed values with a mu of 500 and a sigma of 100
```{r}
rnorm(100, 500, 100)
```

If we want to use the standard normal distrubution, we can omit the $\mu$ and the $\sigma$, and just give n:
```{r}
rnorm(100)
```



