---
title: "4_Vectorized_Functions"
author: "Sophia Barrett"
date: "4/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A **vector** is a one-dimensional array; a collection of items of the same data type.  A **vectorized function** will operate or apply the function to all elements of the vector.  Most of R's functions are vectorized

For example, let's make a vector c:
```{r}
x <- c(1, 2, 3, 4, 5)
x
```
If we square x, we are actually squaring each element in vector x
```{r}
x**2
```
This is the same as writing
```{r}
c(1, 2, 3, 4)**2
```

Adding 2 to x will add 2 to each element in vector x:
```{r}
x + 2
```
we can multiply each element in vector x by 5 and save it to variable y
```{r}
y <- x * 5
y
```

```{r}
z <- x / 100
z
```

We can get the **sum** of the elements in a vector.  The elements of vector x are 1, 2, 3, and 4.  So the sum is 1 + 2 + 3 + 4 = 10:
```{r}
sum(x)
```
We can get the **product** of the elements in vector x (1 * 2 * 3 * 4 = 24)
```{r}
prod(x)
```
The lowest value (element) in vector x:
```{r}
min(x)
```
The highest value (element) in vector x:
```{r}
max(x)
```
We can get the **cumulative sum** for the elements in vector x.   
1     = 1  
1 + 2 = 3  
3 + 3 = 6  
6 + 4 = 10  
```{r}
cumsum(x)
```

We get the *cumulative product** in a similar fashion.  
1     = 1  
1 * 2 = 2  
2 * 3 = 6  
6 * 4 = 24  
```{r}
cumprod(x)
```
We can also combine functions (remember to follow the order of operations)
```{r}
sumOfX <- sum(x)
sumOfX
```

```{r}
sumOfXSquared <- sum(x**2)
sumOfXSquared
```
```{r}
sampleSize = length(x)
sampleSize
```
We can use this in a formula to calculate the sum of squares (SS).  Remember, ${SS}$ = $\sum X^2$-$\left( \frac{\sum X}{n} \right)^2$
```{r}
SS = sumOfXSquared - ((sumOfX)**2/sampleSize)
SS
```

We can have R calculate the **sample variance** of x
```{r}
var(x)
```
As well as the sample standard deviation of x
```{r}
sd(x)
```
We can also use the `sqrt()` to take the square root of the variance, since ${S}$ = $\sqrt{variance}$
```{r}
sqrt(var(x))
```
We work from the inner-most set of function to the outer-most.  In this case,   
1.  get the variance of x, `var(x)`  
2.  then take the square root of the variance with the `sqrt()` function, `sqrt(var(x))`  


We can manipulate the variance formula to get the SS from the variance and the degrees of freedom (df):
```{r}
var(x) * (sampleSize - 1)
```
We already calculated measures of variability.  Let's not forget our measures of central tendency.  Here we find the mean of the x:
```{r}
mean(x)
```
and the median of x:
```{r}
median(x)
```

Finally, we can use the `summary()` function to get the five-number summary as well as the mean for quantitative data.
```{r}
summary(x)
```

## Sorting Vectors

Let's say we have a vector m.
```{r}
m <- c(61, 35, 75, 15, 93, 14, 34, 1, 65)
m
```
We can sort these values using the `sort()` function
```{r}
sort(m)
```
As you can see, we sort the vector components in ASCENDING order by default.

Check out the documentaion for sort() to see how we can sort values in descending order
```{r}
?sort
```
According to the documentation, `decreasing = FALSE` is the default. So to sort in decreasing order, we need to change the parameter to read `decreasing = TRUE`.
```{r}
sort(m, decreasing = TRUE)
```
While we have sorted the values, we have not changed the original vector, m
```{r}
m
```
If we want to save our sorts, we need to store the sorted values in a new vector
```{r}
q <- sort(m, decreasing = TRUE)
q
```



